---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# stubble

<!-- badges: start -->
<!-- badges: end -->

stubble helps you generate simple synthetic datasets matching the format of a supplied data frame-like object (including base R data frames, tibbles, data.tables, and lists of vectors). 

The original intended use of stubble is to generate simple test data for analysis projects and R package development.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("bjcairns/stubble")
```
## Example

Here is a very simple example of using stubble on the `iris` dataset.

``` {r seed, echo=FALSE}
set.seed(98723895)
```

```{r example}
library(stubble)
stubblise(iris)
```

For more advanced use, see the "Using stubble" vignette.

## Known issues

* It is intended that stubble will support generation of character columns with unique values, but this is not currently guaranteed.
